---
title: "RDS203_FinalProject"
author: "Tzu-Hsi (Cece) Jen"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r}
lib <- c("dplyr", "tidyr", "ggplot2", "gghighlight", "ggrepel", "gt")

for(pkg in lib){
  if(!require(pkg, character.only = TRUE)){
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}
```

## 1. Result Overview

```{r}
res <- read.csv("./FinalProject_Code/output/results.csv")
head(res, 3)
```

### 1.1 Average performance

```{r}
# Compute average 
avg_df <- res|>
  group_by(strategy)|>
  summarise(cost_avg = mean(cost),
            qaly_avg = mean(qaly),
            mortalityReduction_avg = mean(mortalityReduction))|>
  mutate(start_age = as.numeric(sub("S(\\d+)_(\\d+)_F(\\d+)", "\\1", strategy)),
         stop_age = as.numeric(sub("S(\\d+)_(\\d+)_F(\\d+)", "\\2", strategy)),
         frequency = as.numeric(sub("S(\\d+)_(\\d+)_F(\\d+)", "\\3", strategy)))|>
  mutate(frequency = recode(frequency,
                            "1" = "Annual",
                            "2" = "Biennial"))

avg_cols <- colnames(avg_df)[grep("_avg$", colnames(avg_df))]
avg_df <- avg_df|>
  mutate(across(all_of(avg_cols), ~ round(.x, digits = 2)))

avg_df
```

### 1.2 Best performance

```{r}
# Find best value
cost_best <- avg_df|>
  slice_min(cost_avg)|>
  mutate(category = "Lowest costs")

qaly_best <- avg_df|>
  slice_max(qaly_avg)|>
  mutate(category = "Highest QALYs")

mortalityReduction_best <- avg_df|>
  slice_max(mortalityReduction_avg)|>
  mutate(category = "Greatest CRC mortality reduction")

best_df <- bind_rows(
  cost_best,
  qaly_best,
  mortalityReduction_best
)
best_df
```

```{r}
best_strategy <- best_df$strategy
compare_df <- res|>
  filter(strategy %in% best_strategy)

t.test(cost ~ strategy, data = compare_df)
t.test(qaly ~ strategy, data = compare_df)
```

```{r, out.width='50%'}
best_tbl <- best_df|>
  select(start_age, stop_age, frequency, category, avg_cols)|>
  arrange(category)|>
  
  gt()|>
  cols_label(
    start_age = "Start Age (years)",
    stop_age = "Stop Age (years)",
    frequency = "Frequency",
    category = "Performance Metric",
    
    cost_avg = "Average Cost (USD)",
    qaly_avg = "Average QALYs",
    mortalityReduction_avg = "Average CRC Mortality Reduction (%)"
  )|>
  opt_table_font(font = "Times New Roman")|>
  tab_style(
    style = cell_text(
      weight = "bold",
      size = px(14),
      align = "center"
      ),
    locations = cells_column_labels()
  )|>
  tab_style(
    style = cell_text(
      align = "center",
      size = px(14)
    ),
    locations = cells_body()
  )


best_tbl
gtsave(data = best_tbl, filename = "best_tbl.png", path = "./table")
```

```{r, out.width='50%'}
avg_tbl <- avg_df|>
  select(strategy, start_age, stop_age, frequency, avg_cols)|>
  gt()|>
  
  tab_style(
    style = cell_fill(color = "lightblue"),
    locations = cells_body(
      rows = strategy %in% best_df$strategy)
  )|>
  cols_hide(strategy)|>
  tab_source_note(
    source_note = md(
      "Highlighted rows represent best-performing screening strategies across selected outcome measures."
    )
  )|>
  
  cols_label(
    start_age = "Start Age (years)",
    stop_age = "Stop Age (years)",
    frequency = "Frequency",
    
    cost_avg = "Average Cost (USD)",
    qaly_avg = "Average QALYs",
    mortalityReduction_avg = "Average Mortality Reduction (%)"
  )|>
  
  opt_table_font(font = "Times New Roman")|>
  tab_style(
    style = cell_text(
      weight = "bold",
      size = px(14),
      align = "center"
      ),
    locations = cells_column_labels()
  )|>
  tab_style(
    style = cell_text(
      align = "center",
      size = px(14),
    ),
    locations = cells_body()
  )

avg_tbl
gtsave(data = avg_tbl, filename = "avg_tbl.png", path = "./table")
```

### 1.2 Worst performance

```{r}
# Find best value
cost_worst <- avg_df|>
  slice_max(cost_avg)|>
  mutate(category = "Highest costs")

qaly_worst <- avg_df|>
  slice_min(qaly_avg)|>
  mutate(category = "Lowest QALYs")

mortalityReduction_worst <- avg_df|>
  slice_min(mortalityReduction_avg)|>
  mutate(category = "Worst mortality reduction")

worst_df <- bind_rows(
  cost_worst,
  qaly_worst,
  mortalityReduction_worst
)
worst_df
```

```{r, out.width='50%'}
worst_tbl <- worst_df|>
  select(start_age, stop_age, frequency, category, avg_cols)|>
  arrange(start_age)|>
  gt()|>
  cols_label(
    start_age = "Start Age (years)",
    stop_age = "Stop Age (years)",
    frequency = "Frequency",
    
    category = "Performance Metric",
    cost_avg = "Average Cost (USD)",
    qaly_avg = "Average QALYs",
    mortalityReduction_avg = "Average Mortality Reduction (%)"
  )|>
  opt_table_font(font = "Times New Roman")|>
  tab_style(
    style = cell_text(
      weight = "bold",
      size = px(14),
      align = "center"
      ),
    locations = cells_column_labels()
  )|>
  tab_style(
    style = cell_text(
      align = "center",
      size = px(14)
    ),
    locations = cells_body()
  )


worst_tbl
gtsave(data = worst_tbl, filename = "worst_tbl.png", path = "./table")
```

## 2. Performance Plot

### 2.1 Best strategy

```{r}
best_label <- best_df|>
  dplyr::rename("Category" = "category")

ggplot(avg_df, aes(x = qaly_avg, y = cost_avg))+
  # all points
  geom_point(colour = "grey", size = 2)+
  
  # highlight best points (best_label)
  geom_point(
    data = best_label,
    aes(colour = Category),
    size = 4)+
  geom_text_repel(
    data = best_label,
    aes(label = strategy, colour = Category),
    size = 3,
    show.legend = FALSE,
    max.overlaps = Inf,
    fontface = "bold",
    box.padding = 0.3,
    point.padding = 0.5,
    segment.color = "grey50")+
  
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  theme_minimal()+
  labs(
    x = "QALYs",
    y = "Cost (USD)"
  )

ggsave("./pic/bestStrategy.png")
```

### 2.2 Cost vs QALYs by frequency

```{r}
ggplot(avg_df, aes(x = qaly_avg, y = cost_avg, colour = factor(frequency)))+
  geom_point(size = 3)+
  labs(
    x = "QALYs",
    y = "Cost (USD)",
    colour = "Frequency"
  )+
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  theme_minimal()

ggsave("./pic/cost_vs_qaly_byFrequency.png")
```

### 2.3 Cost vs QALYs by start age

```{r}
ggplot(avg_df, aes(x = qaly_avg, y = cost_avg, colour = factor(start_age)))+
  geom_point(size = 3)+
  labs(
    x = "QALYs",
    y = "Cost (USD)",
    colour = "Start Age (year)"
  )+
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  theme_minimal()

ggsave("./pic/cost_vs_qaly_byStartAge.png")
```

### 2.4 Cost vs QALYs by stop age

```{r}
ggplot(avg_df, aes(x = qaly_avg, y = cost_avg, colour = factor(stop_age)))+
  geom_point(size = 3)+
  labs(
    x = "QALYs",
    y = "Cost (USD)",
    colour = "Stop Age (year)"
  )+
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  theme_minimal()

ggsave("./pic/cost_vs_qaly_byStopAge.png")
```

### 2.5 QALYs by strategy

```{r}
ggplot(avg_df, aes(x = reorder(strategy, qaly_avg), y = qaly_avg))+
  geom_col(aes(fill = qaly_avg == max(qaly_avg)))+
  scale_fill_manual(values = c("steelblue", "#FF6A6A"), guide = "none")+
  coord_flip(ylim = c(37.0, 38.0))+
  geom_text(aes(label = round(qaly_avg, 3)), hjust = -0.2, size = 3)+
  labs(
    x = "Strategy",
    y = "QALYs"
  )+
  theme_minimal()

ggsave("./pic/qaly_by_strategy.png")
```

### 2.6 Cost by strategy

```{r}
ggplot(avg_df, aes(x = reorder(strategy, cost_avg), y = cost_avg))+
  geom_col(aes(fill = cost_avg == max(cost_avg)))+
  scale_fill_manual(values = c("steelblue", "#FF6A6A"),
                    guide = "none")+
  coord_flip(ylim = c(1400, 2600))+
  geom_text(aes(label = round(cost_avg, 3)), hjust = -0.2, size = 3)+
  labs(
    x = "Strategy",
    y = "Cost (USD)"
  )+
  theme_minimal()

ggsave("./pic/cost_by_strategy.png")
```

### 2.7 Mortality reduction by strategy

```{r}
ggplot(avg_df, aes(x = reorder(strategy, mortalityReduction_avg),
                   y = mortalityReduction_avg))+
  geom_col(aes(fill = mortalityReduction_avg == max(mortalityReduction_avg)))+
  scale_fill_manual(values = c("steelblue", "#FF6A6A"),
                    guide = "none")+
  coord_flip()+
  geom_text(aes(label = round(mortalityReduction_avg, 3)), hjust = -0.2, size = 3)+
  labs(
    x = "Strategy",
    y = "CRC Mortality Reduction (%)"
  )+
  theme_minimal()

ggsave("./pic/MRreduction_by_strategy.png")
```

### 2.8 Tradeoff: Cost vs Mortality (+ Frequency & Start Age)

```{r}
ggplot(avg_df, aes(x = mortalityReduction_avg, y = cost_avg, 
                   colour = factor(frequency),
                   shape = factor(start_age)))+
  
  geom_point(size = 3)+
  
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  
  coord_cartesian(clip = "off")+

  labs(
    x = "CRC Mortality Reduction (%)",
    y = "Cost (USD)",
    colour = "Frequency",
    shape = "Start Age (year)"
  )+
  theme_minimal()

ggsave("./pic/cost_MR_freq_startAge.png")
```

### 2.9 Tradeoff: Cost vs Mortality (+ Frequency & Stop Age)

```{r}
ggplot(avg_df, aes(x = mortalityReduction_avg, y = cost_avg, 
                   colour = factor(frequency),
                   shape = factor(stop_age)))+
  
  geom_point(size = 3)+
  
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  
  coord_cartesian(clip = "off")+

  labs(
    x = "CRC Mortality Reduction (%)",
    y = "Cost (USD)",
    colour = "Frequency",
    shape = "Stop Age (year)"
  )+
  theme_minimal()

ggsave("./pic/cost_MR_freq_stopAge.png")
```

### 2.10 Tradeoff: QALYs vs Mortality (+ Frequency & Start Age)

```{r}
ggplot(avg_df, aes(x = mortalityReduction_avg, y = qaly_avg, 
                   colour = factor(frequency),
                   shape = factor(start_age)))+
  
  geom_point(size = 3)+
  
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  
  coord_cartesian(clip = "off")+

  labs(
    x = "CRC Mortality Reduction (%)",
    y = "QALYs",
    colour = "Frequency",
    shape = "Start Age (year)"
  )+
  theme_minimal()

ggsave("./pic/qalys_MR_freq_startAge.png")
```

### 2.11 Tradeoff: QALYs vs Mortality (+ Frequency & Stop Age)

```{r}
ggplot(avg_df, aes(x = mortalityReduction_avg, y = qaly_avg, 
                   colour = factor(frequency),
                   shape = factor(stop_age)))+
  
  geom_point(size = 3)+

  scale_x_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))+
  
  coord_cartesian(clip = "off")+

  labs(
    x = "CRC Mortality Reduction (%)",
    y = "QALYs",
    colour = "Frequency",
    shape = "Stop Age (year)"
  )+
  theme_minimal()

ggsave("./pic/qalys_MR_freq_stopAge.png")
```

## 3. Calibration Plot

```{r}
calibration_trace <- read.csv("./FinalProject_Code/output/calibration_trace.csv")
head(calibration_trace, 3)
```

```{r}
ggplot(calibration_trace, aes(x = iteration)) +

  geom_line(
    aes(
      y = log(currentError + 1e-10),
      colour = "Current error"
    ),
    alpha = 0.5
  ) +

  geom_line(
    aes(
      y = log(bestError + 1e-10),
      colour = "Best error"
    ),
    linewidth = 1
  ) +

  scale_colour_manual(
    values = c(
      "Current error" = "grey60",
      "Best error" = "red"
    )
  ) +

  labs(
    x = "Iteration",
    y = "Log calibration error",
    colour = NULL
  ) +

  theme_minimal()

ggsave("./pic/calibration_trace.png")
```

## 4. Supplementary Table

### 4.1 Model input 

```{r, out.width='50%'}
input_data <- read.csv("./data/model_input.csv")
head(input_data, 3)

input_tbl <- input_data|>
  gt()|>
  cols_label(
    Age_group_years = "Age Group (years)",
    CRC_Incidence_Rate = "CRC Incidence Rate (years)",
    CRC_Mortality_Rate = "CRC Mortality Rate (years)",
    All_Cause_Mortality_Rate = "All-Cause Mortality Rate (years)"
  )|>
  opt_table_font(font = "Times New Roman")|>
  tab_style(
    style = cell_text(
      weight = "bold",
      size = px(14),
      align = "center"
      ),
    locations = cells_column_labels()
  )|>
  tab_style(
    style = cell_text(
      size = px(14),
      align = "center"
    ),
    locations = cells_body()
  )

input_tbl
gtsave(data = input_tbl, filename = "input_tbl.png", path = "./table")
```

### 4.2 Model calibrated parameters

```{r, out.width='50%'}
calibrated_tbl <- data.frame(
  Parameter = c(
    "Optimisation method",
    "Number of iterations",
    "Calibrated parameter",
    "Calibrated detection effect",
    "Target CRC mortality reduction",
    "Model-predicted CRC mortality reduction",
    "Absolute calibration error"
  ),
  
  Value = c(
    "Simulated annealing",
    1000,
    "Detection effect",
    0.836,
    0.35,
    0.349,
    0.001
  )
)|>
  gt()|>
  opt_table_font(font = "Times New Roman")|>
  tab_style(
    style = cell_text(
      weight = "bold",
      size = px(14),
      align = "center"
      ),
    locations = cells_column_labels()
  )|>
  tab_style(
    style = cell_text(
      size = px(14)
    ),
    locations = cells_body()
  )|>
  cols_align(
    align = "center",
    columns = Value
  )

calibrated_tbl
gtsave(calibrated_tbl, filename = "calibrated_tbl.png", path = "./table")
```












